title

### 2024-12-20 docs/index.qmd
checkpoint Dec 20th

# code to recode levels "-99" in factor vars
data <- data |> 
  mutate(across(where(is.factor), ~forcats::fct_recode(., NULL = "-99"))) |> 
  # Reverse response options to show increase in *less concern*
  mutate(q25 = forcats::fct_rev(q25))

@@

## Design and Measures

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve trust in elections and ease election anxieties, a recent survey experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment (n = 650, 50.51%) control (n = 637, 49.49%) conditions[^Demographic breakdown of the sample are included in Appendix B.].

@@

### Trust in Elections Admin

@@

Note that, except for the item assessing voter confidence (i.e., votes counted as voters intend), most other question items from the 2022 SPAE inquire into the voter experience after the election has occurred. The items in this study, however, all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, other survey questionnaires ask similar questions that specifically contrast between a respondent's own local community, their state, and nationwide. This is in addition, or complementary to, items that distinguish between local area officials, state election officials, state government, and the U.S. federal government. The survey of this study, instead, focused on attitudes regarding election workers such as election officials, staff, and volunteers. As such, the items are utilized to comprise a component[^7] I broadly refer to as *trust in elections*.

[^7]: The methodological concerns between the use of (unit weighted) sum scores compared to the use of estimated factor scores for Likert scales composed of multiple Likert items are legion. The use of either computed sum scores or estimated factor scores as a general practice is continuously debated [@mcneish2020; @widaman2022; @mcneish2023; @widaman2024]. The use of either appears to largely depend on the researcher's particular research objective, theoretical suppositions, and necessity. Another common concern is the inappropriate treatment of ordinal variables as interval which levy strong assumptions, e.g., each item contributes equally to the variable being measured and that ordinal response categories are equidistant as though on an interval scale [@wang2013; for contrast, see @robitzsch2020].

@@

-   How confident are you that votes in Maricopa County, AZ will be counted as voters intend in the elections this November?[^8]
-   How confident are you that election officials, their staff, and volunteers in Maricopa County, AZ will do a good job conducting the elections this November?
-   Think about the election staff and volunteers who handle the administration and conduct of elections in Maricopa County, AZ. How committed do you think they will be to making sure the elections held this November are fair and accurate?
-   How confident are you that the voting process will be fair in Maricopa County, AZ?
-   How confident are you that election systems in Maricopa County, AZ will be secure from hacking and other technological threats?

[^8]: Note that confidence, or lack thereof, that votes will be counted as voters intend could easily be construed as an expectation of electoral fraud. It is a given that the votes will be counted, but the question is whether they will be counted *as intended*. Although fraud isn't provided as the reason for why votes wouldn't be counted as intended, it also isn't implied that miscounts would be due to some negligent mistake. A nefarious insinuation might be somewhat implicit, but perhaps only to those who are inclined to suspect as such. A lack of confidence here doesn't necessarily impart the reason for that lack of confidence, which makes it suitable as a survey item to measure the presence of trust in election administration rather than an expectation of fraud.

@@

-   There will be voter fraud, that is, people who are not eligible to vote will vote, or vote more than once
-   Many votes will not actually be counted
-   Many people will show up to vote and be told they are not eligible
-   A foreign country will tamper with the votes cast in this area to change the results
-   Election officials in Maricopa County, Arizona will try to discourage some people from voting

Similar to the items capturing trust in elections, I computed a composite score for each respondent reflecting their expectations of electoral fraud.

## Model and Data Specification

The outcome of interest concerns both public trust in election administration and the expectation of electoral fraud. Because these variables are assessed by multiple Likert items, conventional methods call for combining the items of each set into separate Likert scales, also known as a summated rating scales.
However, many of the items comprising the trust scale directly oppose items capturing expectations of electoral fraud. For instance, one item has the respondent report their confidence that votes will be counted as voter's intend; whereas a contrasting item has the the respondent report their expectation that many votes will not be counted. These two items directly oppose each other, but there's no guarantee that individual's responded as such.

@@

I would expect composite sum or mean scores, and item responses, on these two measures to be inversely correlated among the sample, but not mutually exclusive. Indeed, this is what I find.

I first constructed two multi-item composite scales of both *trust in elections* and *expectation of fraud*. Cronbach's internal consistency reliability coefficient \alpha were estimated for both scales and were adequate (\alpha > 0.8); the trust in elections scale \alpha = 0.93 and the expectation of electoral fraud scale \alpha = 0.86. In other words, the items in each of the scales were intercorrelated well enough to compose a single scale of a (presumed) unidimensional construct. I then plotted separate histograms of scores across the sample to visualize the distribution for each seen in @fig-dist. Note that for both scales, higher values indicate less expectation of electoral fraud and greater trust in elections, respectively.

@@

-   limitations: interpretations of the treatment effect are limited.
-   The sample may not be generalizable to the population.
-   It's hard to say whether merely mentioning veterans is enough, as compared to explicitly naming veterans as the target of particular recruitment efforts.
-   Interpretation is also limited considering that no other particular group, or groups are compared directly against the veteran treatment vignette, i.e., additional vignettes for other comparable groups.
-   Moreover, results are limited by the survey questions, questionnaire design, and experimental stimulus (i.e., vignette) for two reasons. - First, the vignettes and many survey questions ask about Maricopa County, AZ specifically. Adding a specific county in the vignette adds in a factor that cannot be accounted for without additional treatments that eliminate the setting as a potential influence. Moreover, the specificity of the setting adds in even more unexplained error---some people may have attitudes about the county in question, others won't, while others may be muffed to consider a random county in the U.S. they've know nothing about. Adding the county undermines the confidence that treatment effects are solely attributable to veterans to an unknown degree. - Second, questions ask about one's own local community in addition to identical questions which asked about Maricopa County, AZ. Consequently, many questions within the questionnaire were duplicates with that one differences. This isn't uncommon in surveys, however in this case, it lengthened the survey to a degree that likely resulted in a higher drop off rate. More importantly, the quality of responses were likely diminished to some unknown extent. Although it is possible to compare questions asked about Maricopa County, AZ to questions about one's local area, the quality of that comparison is limited by the unknown extent of fatigue induced by answering the same questions twice. It wasn't just that some questions were asked multiple times, almost all of the questions were duplicated; after completing a long series on Maricopa County, participants then answered the same questions about the local community. The rate to which participants dropped from the survey is \_\_\_, which suggests fatigue as an important factor. Survey fatigue is a known issue that should be taken into consideration. Comparison of treatment effects between the MC series and the local area series is undermined by the unknown influence that fatigue would have on response choices. Any differences couldn't confidently be attributed to "my area" vs Maricopa county. We also can't determine whether treatment effects are sustained when comparing the Maricopa County series to the local area series asked subsequently. We would have to assume the setting as irrelevant, which we can't reasonably do.

@@

::: {.content-visible when-format="html" unless-format="pdf"}
```{r}
#| label: tbl-desc
#| tbl-cap: "Description of Sample Demographics"
#| tbl-cap-location: top

# quick, informative, easy table of sample demographics w/ some other questions.
tbl.desc <- data |>  
  labelled::set_variable_labels(
    age_cat = "Age group, in years"
  ) |> 
  gtsummary::tbl_summary(
    by = qset,
    include = c(
    qset,
    age_cat, 
    gender_3cat, 
    race, 
    educ_4cat, 
    partyid_3cat,
    ideo,
    ideolean,
    voted2020,
    choice2020, 
    voteintent,
    milserv1,
    milserv2,
    milservfam),
    statistic = list(gtsummary::all_categorical()~ "{n} ({p}%)"),
    type = list(milserv1 ~ "categorical",

                milserv2 ~ "categorical",
                milservfam ~"categorical"), 
    missing = "no") |> 
  gtsummary::add_n() |> 
  gtsummary::add_overall() |> 
  gtsummary::modify_header(label = "**Variable**") |> 
  gtsummary::bold_labels()

gtsummary::as_gt(tbl.desc) |>  
  gt::tab_footnote(
    footnote = "'true' Independents do not identify nor 'lean' toward either political party",
    placement = "left"
    ) |>  
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

@@

::: {.content-hidden unless-format="pdf"}
```{r}
#| label: tbl-desc-pdf
#| tbl-cap: "Description of Sample Demographics"
#| tbl-cap-location: top

frqs <- data |> 
  select(group, qset, age_cat, gender_3cat, educ_4cat, race, partyid_3cat) |> 
  map(count_prop)

demog_df <- purrr::reduce(frqs, dplyr::full_join)

kableExtra::kbl(demog_df,
                booktabs = T, 
                digits = 2,
                col.names = c("Name", "n", "Pct")) |> 
  kableExtra::pack_rows("Experiment Condition", 1, 2) |>
  kableExtra::pack_rows("Question Set", 3, 4) |>
  kableExtra::pack_rows("Age", 5, 12) |> 
  kableExtra::pack_rows("Gender", 13, 15) |> 
  kableExtra::pack_rows("Education", 16, 19) |> 
  kableExtra::pack_rows("Race", 20, 24) |> 
  kableExtra::pack_rows("Party ID", 25, 27) |> 
  kableExtra::kable_styling(
    latex_options = c("striped", "scale_down"),
    full_width = F,
    font_size = 8) |> 
  kableExtra::column_spec(1, width = "20em")

@@

[^1]: I adopt a distinction made between *trustworthiness* and *trust* in elections [@stewart2022]. The "*worthiness*" of one's trust in the conduct and administration of elections is based on the extent that outcomes of an election reasonably follow the rules prescribed and can be adjudicated as such. "If the process is conducted fairly and competently, and the results are determined by the actions of voters...we can call this a trustworthy election" [@stewart2022, 237]. *Trustworthiness* is distinguished from *trust* as a psychological construct, "...the conclusion reached by the public about the functioning of the process" [@stewart2022, 237]. To put in other terms, trustworthiness is built by the structure, procedures, and practices of the institution, in this case election administration. The public's trust in elections, however, is amendable to an indefinite number of factors that may be seemingly unrelated to the formal procedures of election administration. To illustrate, my car is trustworthy because I've known it to function quite well as a car should; it passes whatever criteria upon inspection, its sufficiently fueled, and is in generally good working order. I have every reason to deem it worthy of my trust. However, I don't trust my car because I am sure it is haunted.

@@

[^4]: The cited study defined "insurrectionist sentiments" as, "...willingness to support violent efforts to overturn the results of an election in favor of another, unelected political leader." [@pape2024, 7]. It should be noted, however, that despite the general definition, the operative items used to measure such sentiments ask about Donald Trump specifically, not an "unelected political leader".

@@

>>>>

@@

## Design and Survey Measures

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve trust in elections and ease election anxieties, a recent survey experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment (n = 650) control (n = 637) conditions[^4].

@@

### Trust in Elections

@@

Note that, except for the item assessing voter confidence (i.e., votes counted as voters intend), most other question items from the 2022 SPAE inquire into the voter experience after the election has occurred. The items in this study, however, all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, other survey questionnaires ask similar questions that specifically contrast between a respondent's own local community, their state, and nationwide. This is in addition, or complementary to, items that distinguish between local area officials, state election officials, state government, and the U.S. federal government. The survey of this study, instead, focused on attitudes regarding election workers such as election officials, staff, and volunteers. As such, the items are utilized to comprise a component[^6] I broadly refer to as *trust in elections*.

@@

::: {#tbl-1}
| **Trust in Elections** |
|---|
| How confident are you that votes in Maricopa County, AZ will be counted as voters intend in the elections this November?[^7] |
| How confident are you that election officials, their staff, and volunteers in Maricopa County, AZ will do a good job conducting the elections this November? |
| Think about the election staff and volunteers who handle the administration and conduct of elections in Maricopa County, AZ. How committed do you think they will be to making sure the elections held this November are fair and accurate? |
| How confident are you that the voting process will be fair in Maricopa County, AZ? |
| How confident are you that election systems in Maricopa County, AZ will be secure from hacking and other technological threats? |
:::

@@

::: {#tbl-2}
| **Expectation of Electoral Fraud** |
|---|
| There will be voter fraud, that is, people who are not eligible to vote will vote, or vote more than once |
| Many votes will not actually be counted |
| Many people will show up to vote and be told they are not eligible |
| A foreign country will tamper with the votes cast in this area to change the results |
| Election officials in Maricopa County, Arizona will try to discourage some people from voting |
:::

Similar to the items capturing trust in elections (@tbl-1), I computed a composite score for each respondent reflecting their expectations of electoral fraud.

## Data Specification

The outcome of interest concerns both public trust in election administration and the expectation of electoral fraud. Because these variables are assessed by multiple Likert items, conventional methods call for combining the items of each set into separate Likert scales, also known as a summated rating scales.
However, many of the items comprising the trust scale directly oppose items capturing expectations of electoral fraud. For instance, one item has the respondent report their confidence that votes will be counted as voter's intend; whereas a contrasting item has the the respondent report their expectation that many votes will not be counted. These two items directly oppose each other, but there's no guarantee that individual responded as such.

@@

I would expect composite sum or mean scores, and item responses, on these two measures to be inversely correlated among the sample, but not mutually exclusive. Indeed, this is what I find. Upon construction of the two multi-item composite scales of *trust in elections* and *expectation of fraud*, internal consistency reliability coefficient $\alpha$ were estimated for both and found to be adequate ($\alpha$ > 0.8); the trust in elections scale $\alpha$ = 0.93 and the expectation of electoral fraud scale $\alpha$ = 0.86. In other words, the items in each of the scales were intercorrelated well enough to compose a single scale of a (presumed) unidimensional construct. I then plotted separate histograms of scores across the sample to visualize the distribution for each seen in @fig-dist. Note that for both scales, higher values indicate less expectation of electoral fraud and greater trust in elections, respectively.

@@

* limitations: interpretations of the treatment effect are limited.
  + The sample may not be generalizable to the population.
  + It's hard to say whether merely mentioning veterans is enough, as compared to explicitly naming veterans as the target of particular recruitment efforts.
* Interpretation is also limited considering that no other particular group, or groups are compared directly against the veteran treatment vignette, i.e., additional vignettes for other comparable groups.
* Moreover, results are limited by the survey questions, questionnaire design, and experimental stimulus (i.e., vignette) for two reasons. - First, the vignettes and many survey questions ask about Maricopa County, AZ specifically. Adding a specific county in the vignette adds in a factor that cannot be accounted for without additional treatments that eliminate the setting as a potential influence. Moreover, the specificity of the setting adds in even more unexplained error---some people may have attitudes about the county in question, others won't, while others may be muffed to consider a random county in the U.S. they've know nothing about. Adding the county undermines the confidence that treatment effects are solely attributable to veterans to an unknown degree. - Second, questions ask about one's own local community in addition to identical questions which asked about Maricopa County, AZ. Consequently, many questions within the questionnaire were duplicates with that one differences. This isn't uncommon in surveys, however in this case, it lengthened the survey to a degree that likely resulted in a higher drop off rate. More importantly, the quality of responses were likely diminished to some unknown extent. Although it is possible to compare questions asked about Maricopa County, AZ to questions about one's local area, the quality of that comparison is limited by the unknown extent of fatigue induced by answering the same questions twice. It wasn't just that some questions were asked multiple times, almost all of the questions were duplicated; after completing a long series on Maricopa County, participants then answered the same questions about the local community. The rate to which participants dropped from the survey is \_\_\_, which suggests fatigue as an important factor. Survey fatigue is a known issue that should be taken into consideration. Comparison of treatment effects between the MC series and the local area series is undermined by the unknown influence that fatigue would have on response choices. Any differences couldn't confidently be attributed to "my area" vs Maricopa county. We also can't determine whether treatment effects are sustained when comparing the Maricopa County series to the local area series asked subsequently. We would have to assume the setting as irrelevant, which we can't reasonably do.

@@

::: {.content-visible when-format="html" unless-format="pdf"}
```{r}
#| label: tbl-desc
#| tbl-cap: "Demographics and Balance"
#| tbl-cap-location: top

# quick, informative, easy table of sample demographics w/ some other questions.
# best method for html document
tbl.desc <- data |>  
  labelled::set_variable_labels(
    age_4cat = "Age group, in years",
    gender_3cat = "Gender",
    race2 = "Race/Ethnicity",
    educ_4cat = "Education",
    partyid_3cat = "Party ID"
  ) |> 
  gtsummary::tbl_summary(
    by = group,
    include = c(
    age_4cat, 
    gender_3cat, 
    race2, 
    educ_4cat, 
    partyid_3cat),
    statistic = list(gtsummary::all_categorical()~ "{n} ({p}%)"), 
    missing = "ifany",
    missing_text = "NA") |> 
  gtsummary::add_n() |> 
  gtsummary::add_overall() |> 
  gtsummary::modify_header(label = "**Variable**") |> 
  gtsummary::bold_labels()

gtsummary::as_gt(tbl.desc) |>  
  gt::tab_footnote(
    footnote = "Independents do not identify nor 'lean' toward either political party",
    placement = "left"
    ) |>  
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

@@

::: {.content-hidden unless-format="pdf"}
```{r}
#| label: tbl-desc-pdf
#| tbl-cap: "Demographics and Balance"
#| tbl-cap-location: top

gtsummary::as_gt(tbl.desc) |>  
  gt::tab_footnote(
    footnote = "Independents do not identify nor 'lean' toward either political party",
    placement = "left"
    ) |>  
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

@@

[^1]: I adopt a distinction made between *trustworthiness* and *trust* in elections [@stewart2022]. The "*worthiness*" of one's trust in the conduct and administration of elections is based on the extent that outcomes of an election reasonably follow the rules prescribed and can be adjudicated as such. To put it briefly, trustworthiness is built by the structure, procedures, and practices of the institution, in this case election administration. The public's trust in elections, however, is amendable to an indefinite number of factors that may be unrelated to the formal structure or procedures of election administration. To illustrate, I can recognize that my car is trustworthy prior to ever driving it because it is structurally sound; it passes whatever criteria upon inspection, it is sufficiently fueled, and appears to be in generally good working order. I have every reason to deem it *worthy of my trust*. However, I don't trust my car because I am sure it is haunted.

@@

[^4]: Demographic breakdown of the sample are included in Appendix B.

@@

[^6]: The methodological concerns between the use of (unit weighted) sum scores compared to the use of estimated factor scores for Likert scales composed of multiple Likert items are legion. The use of either computed sum scores or estimated factor scores as a general practice is continuously debated [@mcneish2020; @widaman2022; @mcneish2023; @widaman2024]. The use of either appears to largely depend on the researcher's particular research objective, theoretical suppositions, and necessity. Another common concern is the inappropriate treatment of ordinal variables as interval which levy strong assumptions, e.g., each item contributes equally to the variable being measured and that ordinal response categories are equidistant as though on an interval scale [@wang2013; for contrast, see @robitzsch2020].

[^7]: Note a lack of confidence that votes will be counted as voters intend does not imply an expectation of electoral fraud. The reason for one's lack of confidence is uncertain, which makes it suitable as a survey item to measure the presence of trust in election administration rather than an expectation of fraud.
